---
title: "synthetic_control_exploration"
output: 
  html_document:
    toc: true
    toc_float: true
---
<style>
/* Table of Contents Styles */
#TOC {
  color: #333333;
  font-family: 'Arial', sans-serif; 
  font-size: 16px; 
  border-color: #333333;
}

/* Header Styles */
#header {
  color: #B22222; 
  background-color: #F9F9F9; 
  opacity: 0.9; 
  font-family: 'Arial', sans-serif;
  font-size: 24px;
  padding: 15px; 
}

/* Body Styles */
body {
  color: #000000; 
  font-family: 'Arial', sans-serif;
  background-color: #FFFFFF; 
  line-height: 1.6; 
}

/*  Text Styles */
pre {
  color: #4D4D4D; 
  background-color: #F0F8FF; 
  border: 1px solid #B0C4DE; 
  padding: 10px; 
}

/* Header Styles */
h1, h2, h3 {
  color: #003366;
  font-family: 'Arial', sans-serif;
}

/* Link Styles */
a {
  color: #003366; 
  text-decoration: none; 
}

a:hover {
  text-decoration: underline; 
}
</style>

# Introduction

One really interesting estimator I never went through at university in any real depth was **Synthetic Control**. This is a great shame because it is a really exciting and honestly quite old estimator now *(and when I was studying!)*.

So I thought it might be nice to add a brief *(yet quite in depth)* exploration of this estimator! 

The original and commonly used package in `R` is the `Synth` package by: [insert citation].





```{r setup, message=FALSE}
library(data.table)
library(tidyverse)
library(Synth)
library(haven)
library(optimx)
library(kernlab)

options(scipen = 999)

```

## Synthetic Control Summary

To explore the estimator I will use the famous example from ... 




```{r simple example}
co2 <- setDT(read_dta("/home/michael/Projects/essays/swedish_cp_replication/carbontax_fullsample_data.dta"))

#select countries and filter for years
co2_clean <- co2[country %in% c("Sweden",
                  "Australia",
                  "Belgium",
                  "Canada",
                  "Denmark",
                  "France",
                  "Greece",
                  "Iceland",
                  "Japan",
                  "New Zealand",
                  "Poland",
                  "Portugal",
                  "Spain",
                  "Switzerland",
                  "United States") & 
                   year > 1959 &
                   year < 2006]

# fix id
co2_clean[, Countryno := .GRP, by = .(country)]

# clean up environment
rm(co2)

# data prep object - we will explore all this intense looking stuff in depth later :)
D <- dataprep(foo = co2_clean,
              predictors = c("GDP_per_capita", 
                             "gas_cons_capita", 
                             "vehicles_capita", 
                             "urban_pop"),
              predictors.op = "mean",
              time.predictors.prior = 1980:1989,
              special.predictors = list(
                  list("CO2_transport_capita" , 1989 , "mean"),
                  list("CO2_transport_capita" , 1980 , "mean"),
                  list("CO2_transport_capita" , 1970 , "mean")
              ),
              dependent = "CO2_transport_capita",
              unit.variable = "Countryno",
              unit.names.variable = "country",
              time.variable = "year",
              treatment.identifier = 13,
              controls.identifier = c(1:12,14:15),
              time.optimize.ssr = 1960:1989,
              time.plot = 1960:2005)

# this is the wrapper function that calculates all the nice stuff for our synthetic control weights
synth_out <- synth(data.prep.obj = D, method = 'L-BFGS-B')

# fitted weight vectors (v for the predictors weights and w for the unit weights)
synth_out$solution.v
synth_out$solution.w

# simple effects calcs
effects <- D$Y1plot - (D$Y0plot %*% synth_out$solution.w)
sum(effects[32:length(effects)])
mean(effects[32:length(effects)])

# percentage changes
effects_prct <- (D$Y1plot - (D$Y0plot %*% synth_out$solution.w))/2.309894
sum(effects_prct[32:length(effects)])
mean(effects_prct[32:length(effects)])

# tables - creates some nice objects for reporting
synth_tables <- synth.tab(dataprep.res = D, synth.res = synth_out)

## table for weights w
synth_tables$tab.w

## special predictors (X matrices)
synth_tables$tab.pred
synth_tables$tab.v

# plots from ggplot
## path plot
data_plot <- data.table(year = 1960:2005,
                        co2_actual = D$Y1plot,
                        co2_synth = D$Y0plot %*% synth_out$solution.w,
                        effects = D$Y1plot - (D$Y0plot %*% synth_out$solution.w))

names(data_plot) <-  c("year", "co2_actual", "co2_synth", "effects")

data_plot %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = co2_actual, color = "Actual")) +
  geom_line(aes(y = co2_synth, color = "Synthetic")) +
  labs(title = "Path Plot of Actual vs Synthetic Outcomes",
       x = "Year",
       y = "Metric tons per capita (CO2 from transport)") +
  scale_color_manual(name = "Outcome", values = c("Actual" = "cornflowerblue", "Synthetic" = "blue")) +
  theme_classic()+
  theme(legend.position = "top")+
  geom_vline(xintercept = 1990, linetype = 'dashed', color = "black")


## gap plot
data_plot %>% 
  ggplot(aes(x = year, color)) +
  geom_line(aes(y = effects), color = "blue") +
  labs(title = "Gap plot of calculated effect",
       x = "Year",
       y = "Gap in metric tons per capita (CO2 from transport)") +
  theme_classic()+
  theme(legend.position = "top")+
  geom_hline(yintercept = 0, linetype = 'dashed', color = "black")+
  geom_vline(xintercept = 1990, linetype = 'dashed', color = "black")


# placebo test for loop
## initialise data
store <- matrix(NA,length(1960:2005),15)
colnames(store) <- unique(co2_clean$country)

# run loop
for(iter in 1:15){
 dataprep.out <-
              dataprep(foo = co2_clean,
                       predictors = c("GDP_per_capita" ,
                                      "gas_cons_capita" , 
                                      "vehicles_capita" ,
                                      "urban_pop") ,
                       predictors.op = "mean" ,
                       time.predictors.prior = 1980:1989 ,
                       special.predictors = list(
                         list("CO2_transport_capita" , 1989 , "mean"),
                         list("CO2_transport_capita" , 1980 , "mean") ,
                         list("CO2_transport_capita" , 1970 , "mean")
                         ),
                       dependent = "CO2_transport_capita",
                       unit.variable = "Countryno",
                       unit.names.variable = "country",
                       time.variable = "year",
                       treatment.identifier = iter, # note here the loop changes the treated unit for each loop
                       controls.identifier = c(1:15)[-iter],
                       time.optimize.ssr = 1960:1989,
                       time.plot = 1960:2005
                       )

# run synth
synth.out <- synth(data.prep.obj = dataprep.out, method = 'L-BFGS-B')

# store gaps
store[,iter] <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)
} # end of placebo loop

data <- store
rownames(data) <- 1960:2005

# Set bounds
gap.start     <- 1
gap.end       <- nrow(data)
years         <- 1960:2005
gap.end.pre  <- which(rownames(data)=="1989")

#  MSPE Pre-Treatment
mse  <- apply(data[ gap.start:gap.end.pre,]^2,2,mean)
sweden.mse <- as.numeric(mse[13])
# Exclude MSPE > 20 
data <- data[,mse<20*sweden.mse]
Cex.set <- 1

data_long <- data.table::melt(as.data.table(data, keep.rownames = TRUE),
                              id.vars = "rn",
                              variable.name = "Country",
                              value.name = "Gap")

data_long$Year <- as.integer(data_long$rn)

data_long %>% 
  ggplot(aes(x = Year, y = Gap, group = Country)) +
  geom_line(data = subset(data_long, Country != "Sweden"), 
            aes(color = "Control Countries"), color = "cornflowerblue", alpha = 0.4) +
  geom_line(data = subset(data_long, Country == "Sweden"), 
            aes(color = "Sweden"), color = "blue", linewidth = 1) +
  geom_vline(xintercept = 1990, linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 1) +
  scale_x_continuous(breaks = seq(1960, 2005, 5)) +
  scale_y_continuous(limits = c(-0.5, 0.5)) +
  labs(x = "Year", y = "Gap in metric tons per capita (CO2 from transport)",
       title = "Gap in CO2 Emissions from Transport per Capita",
       subtitle = "Comparison between Sweden and Control Countries (Placebo)") +
  theme_classic() +
  theme(legend.position = "bottomleft") +
  annotate("text", x = 1983, y = -0.5, label = "Carbon Tax", size = 4) +
  annotate("segment", x = 1987, xend = 1989, y = -0.5, yend = -0.5, color = "black", arrow = arrow(length = unit(0.1, "inches"))) +
  guides(color = guide_legend(override.aes = list(linetype = c(1, 1), color = c("black", "gray"), size = c(1, 0.5))))

# Create a data frame for the weights
data.frame(country = synth_tables$tab.w$unit.names,
           Weight = synth_tables$tab.w$w.weights) %>% 
  ggplot(aes(x = reorder(country, -Weight), y = Weight)) +
  geom_bar(stat = "identity", fill = "cornflowerblue", color = "blue", alpha = 0.4) +
  labs(title = "Synthetic Control Unit Weights",
       x = "Countries",
       y = "Weights") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

rm(list=setdiff(ls(), "co2_clean"))

```

## A deep dive into Synthetic Control

Now that we've seen a simple application using `Synth` lets dig into the code a bit deeper to get a better idea at what is going on underneath the hood! 

Firstly, let's look at that `dataprep` object and what it does:

```{r parameters}
# below are the key arguments used in estimating the synthetic control weights for the data prep object

# data prep objects parameters
foo <-  as.data.frame(co2_clean) # this is the data.frame or table your data is stored in. In the above example this is the co2 data

# these are the predictors that form (part of the) X matrices used in the optimisation problem. This is a list of the column names as strings
predictors <-  c("GDP_per_capita", 
                             "gas_cons_capita", 
                             "vehicles_capita", 
                             "urban_pop") 

predictors.op <-  "mean" # the operation applie to the treated X matrix (i.e. mean or median etc - mean is usually sufficient) 

# these are the (special) predictors that form (part of the) X matrices used in the optimisation problem. This is a list of a list of the lags for the output variable we want to include.
special.predictors <-  list( 
                  list("CO2_transport_capita" , 1989 , "mean"),
                  list("CO2_transport_capita" , 1980 , "mean"),
                  list("CO2_transport_capita" , 1970 , "mean")
              ) 

# this is the variable of interest
dependent <-  "CO2_transport_capita"

# this is the column name identifying the unit dimension i.e. countries numebrs "Countryno"
# the code turns this into a column position index, so we can skip that part and just get set it as the first column 
unit.variable <-  1

# unit varibale names is the names we want to include in the final outputs i.e. the countries actual names and not just an index number
# the code also turns this into a column position index, so we can skip that part and just get set it as the 2nd column (country)
unit.names.variable <-  2

# the column that his the time dimension in our panel
# Similarly, the code turns this into a column position index, so we can skip that part and just get set it as the 3 column (year)
time.variable <-  3

# this is a arguement that identifies which of the `unit.variable` is the treated unit. For out case it is unit 13
treatment.identifier  <- 13

# this is a list identifying the remaining units that are controls (only one unit can be treated in this example)
controls.identifier <-  c(1:12,14:15)

# a list identifying the pretreatment periods over which the special predictors should be averaged
time.predictors.prior <-  1980:1989

# a list identifying the pretreatment periods over which the loss function is minimised
time.optimize.ssr <-  1960:1989

# this sets up the time we want to plot the data. Not important for our case tbh.
# if nothing is provided, this will default to `time.optimize.ssr`. But in our case we will set this to the whole period.
time.plot <-  1960:2005



```

Now that we've set up all the key arguements, let's explore the actual transformations.

```{r transformations}

# Firstly, lets identify the treated id country
# the variable for this is 
treatment.identifier.name <- co2_clean[Countryno == treatment.identifier, unique(country)]

# next let's identify the control variables:
controls.identifier.name <- co2_clean[Countryno != treatment.identifier, unique(country)]

# next we sort by unit and then time
setorder(foo, Countryno, year)

# next the code extracts the rows relating to treatment and control units
treatment.rows <- which(foo[,unit.variable] %in% treatment.identifier)
control.rows   <- which(foo[,unit.variable] %in% controls.identifier)

# next we want to extract the time time rows for use later
time.predictors.prior.rows <- which(foo[,time.variable] %in% time.predictors.prior)
time.optimize.ssr.rows <- which(foo[,time.variable] %in% time.optimize.ssr)

# okay now we have the key inputs, we can start building the X matrices
## we will start with the treated matrix X_1
## basically, we just look for the intersection with the treated.rows we got above and return all the predictors we listed above :)
X1 <- data.frame(foo[intersect(treatment.rows, time.predictors.prior.rows), predictors])

# next we calculating the means of X1 

# now we can add the column and row names (helps later on)
colnames(X1) <- predictors
rownames(X1) <- time.predictors.prior

# now we aggregate X taking its mean
X1 <- as.matrix(apply(X1, 2, paste(predictors.op), na.rm = TRUE))
X1

# now we can add the column and row names (yet inverted)
rownames(X1) <- predictors
colnames(X1) <- treatment.identifier

# now we repeat for the control population
X0 <-  data.frame(foo[intersect(control.rows, time.predictors.prior.rows), c(predictors, "Countryno")] )

# now before aggregating, we split into the different units then apply the mean
X0 <- split(X0, X0[,dim(X0)[2]])
X0 <- sapply(X0, apply, 2, mean, na.rm = TRUE, simplify = TRUE)
X0 <- as.matrix(X0[-dim(X0)[1],])

X0

# next we have to add the `special.predictors` to the matrices above. 
# for our approache these are some lagged versions of the outcome variable (means) 
# `Synth` uses a wrapper function for this, but we can just explore a simplified version of it

# variables to help with the below forloop
tr.numb <- treatment.identifier
co.numb  <- controls.identifier
unit.var <-  unit.variable #"Countryno"
time.var <-  time.variable
foo.object <- foo
X0.inner <- X0
X1.inner <- X1
          
          
# however we will need to apply the method to each special predictor i.e. a for-loop
for(i in 1:length(special.predictors)){
          list.object <- special.predictors[[i]]
          
          
          # extract treated and contrl row identifiers
          special.units.tr <- which(is.element(foo.object[,unit.var], tr.numb) == TRUE)
          special.units.co <- which(is.element(foo.object[,unit.var], co.numb) == TRUE)
          
          # extract special time periods for transformation
          special.times <- which(is.element(foo.object[,time.var], list.object[[2]]) == TRUE)
          
          # extract special predictor column 
          # if(mode(list.object[[1]]) == "character"){
          #   list.object[[1]] <- which(names(foo.object) == list.object[[1]])
          #   }
          # 
          # create names - this might look confusing but it is just pasting together strings from the special predictors list provided :)
          name.predictor <- paste("special.", list.object[[1]], ".", paste(list.object[[2]], sep = ""), sep = "")
          
          # now we can create X1
          X1.special <- as.matrix(foo.object[intersect(special.units.tr, special.times), list.object[[1]]])
          
          # name rows and columns again 
          rownames(X1.special) <- foo.object[intersect(special.units.tr, special.times), time.var]
          colnames(X1.special) <- tr.numb
          
          # aggregate if only one time period is there (this is probably unecessary for our example)
          if(length(list.object[[2]]) > 1){
            X1.special <- apply(X1.special,
                            2,
                            paste(list.object[[3]]),
                            na.rm = TRUE)
            }
          
          # now we can combine it to the existing X1
          X1.special <- t(as.matrix(X1.special))
          X1.inner <- rbind(X1.inner, X1.special)
          rownames(X1.inner)[nrow(X1.inner)] <- name.predictor
          
          
          # now we basically repeat this for the controls
          X0.special <- as.matrix(foo.object[intersect(special.units.co, special.times), c(list.object[[1]])])
          
          X0.special <- matrix(X0.special[,1], byrow = FALSE, ncol = length(co.numb))
          
          ## deinfe the columns and rows
          row.names(X0.special) <- foo.object[intersect(special.units.tr, special.times), time.var]
          colnames(X0.special) <- co.numb
          
           if(length(list.object[[2]]) > 1){
             X0.special <- apply(
                            X0.special,
                            2,
                            paste(list.object[[3]]),
                            na.rm = TRUE
                            )
             X0.special <- t(as.matrix(X0.special))
           }
          
          X0.inner <- rbind(X0.inner,X0.special)
          rownames(X0.inner)[nrow(X0.inner)] <- name.predictor
          
          X1 <- X1.inner
          X0 <- X0.inner

           # now go and built matrices
           # spf <- spec.pred.func(list.object = special.predictors[[i]],
           #                        tr.numb = treatment.identifier,
           #                        co.numb = controls.identifier,
           #                        unit.var = unit.variable,
           #                        time.var = time.variable,
           #                        foo.object = foo,
           #                        X0.inner = X0,
           #                        X1.inner = X1
           #                        )
           # 
           #  X0 <- spf[[1]]
           #  X1 <- spf[[2]]

          }

# clear up environ after loop
rm(X0.special, X1.special,  X0.inner, X1.inner, tr.numb, co.numb , unit.var, time.var, foo.object)
gc()

# great we now have the general X matrices! Now we can go onto the Z matrices :)
Z1 <- as.matrix(foo[intersect(treatment.rows,time.optimize.ssr.rows), dependent])

## set up row and col names again 
rownames(Z1) <- time.optimize.ssr
colnames(Z1) <- treatment.identifier

# now for Z0 (controls) - note it has to be sent wide such that each column refers to one treated unit
# Z0 <- as.matrix(foo[intersect(control.rows,time.optimize.ssr.rows), c(dependent)])

Z0 <- matrix(as.matrix(foo[intersect(control.rows,time.optimize.ssr.rows), c(dependent)]),
             byrow = FALSE,
             nrow = length(time.optimize.ssr),
             ncol = length(controls.identifier)
             )
                 
colnames(Z0)<- controls.identifier
rownames(Z0)<- time.optimize.ssr

# now we have all the key elements for the optimisation problem, but one extra thing the function does
# is return the plot time series for the treated unit, plus the matrices for the control units to create the counterfactual
# this is basically more of the same above but now we're using `time.variable` instead of `tine.optimize.ssr` or `time.predictors.prior` 
# hopefully these transformations don't need to much more explanation

# treated time pliot
time.plot.rows <- which(is.element(foo[,time.variable], time.plot ) == TRUE)

Y1plot <- as.matrix(foo[intersect(treatment.rows, time.plot.rows), dependent])
rownames(Y1plot) <- time.plot
colnames(Y1plot) <- treatment.identifier

Y0plot  <- matrix(as.matrix(foo[intersect(control.rows, time.plot.rows), c(dependent)]),
                  byrow = FALSE,
                  nrow = length(time.plot),
                  ncol = length(controls.identifier))
rownames(Y0plot) <- time.plot
colnames(Y0plot) <- controls.identifier

# the last little thing the data prep function does that's helpful is store all this in a nice list to call much easier later on

# this returns a data.frame matchign the unit names and their numbers so you can easily track which number relates to which unit
names.and.numbers <- data.frame(c(treatment.identifier.name,controls.identifier.name), c(treatment.identifier,controls.identifier))
names(names.and.numbers) <- c("unit.names", "unit.numbers")

output <- list(X0 = X0,
               X1 = X1,
               Z0 = Z0,
               Z1 = Z1,
               Y0plot = Y0plot,
               Y1plot = Y1plot,
               names.and.numbers = names.and.numbers)

```


Okay great, we've now for all the data structures we need to estimate $W$. We have $Z_1$, $Z_0$, $X_1$ and $X_0$. We also have the matrices to calculate the counterfactual. 

But before, we do that we have to solve the optimisation problem! 

There are a lot more checks and balances in the actual source code to make sure you don't mess anything up. But for our purposes of exploration, the above will suffice :)

```{r optimisation function}

#View(fn.V)

test_fn.V <-function(
           variables.v = stop("variables.v missing"),
           X0.scaled = stop("X0.scaled missing"),
           X1.scaled = stop("X1.scaled missing"),
           Z0 = stop("Z0 missing"),
           Z1 = stop("Z1 missing"),
           margin.ipop = 0.0005,
           sigf.ipop = 5,
           bound.ipop = 10
           ){
    # rescale par
    V <- diag(x=as.numeric(abs(variables.v)/sum(abs(variables.v))),
              nrow=length(variables.v),
              ncol=length(variables.v))
    
    # set up QP problem
    H <- t(X0.scaled) %*% V %*% (X0.scaled)
    a <- X1.scaled
    c <- -1*c(t(a) %*% V %*% (X0.scaled) )
    A <- t(rep(1, length(c)))
    b <- 1
    l <- rep(0, length(c))
    u <- rep(1, length(c))
    r <- 0
    
    # run solver and obtain w weights
    res <- ipop(c = c, H = H, A = A, b = b, l = l, u = u, r = r,
                  bound = bound.ipop, margin = margin.ipop,
                  maxiter = 1000, sigf = sigf.ipop)
    solution.w <- as.matrix(primal(res))
            
    # compute losses
    loss.w <- as.numeric(t(X1.scaled - X0.scaled %*% solution.w) %*% (V) %*% (X1.scaled - X0.scaled %*% solution.w))

    loss.v <- as.numeric(t(Z1 - Z0 %*% solution.w) %*% (Z1 - Z0 %*% solution.w ))
    loss.v <- loss.v/nrow(Z0)

    return(invisible(loss.v))
  }

```



So, let's explore how we can solve the optimisation problem. 

```{r optimisation problem}

# The first thing we do is normalise X by the variance - this returns some scaled matrices to us (scaled by variance of each predictor)
nvarsV <- dim(X0)[1]
big.dataframe <- cbind(X0, X1)
divisor <- sqrt(apply(big.dataframe, 1, var))

scaled.matrix <- t(t(big.dataframe) %*% ( 1/(divisor) * diag(rep(dim(big.dataframe)[1], 1)) )) # last bit is the identify matrix such that it becomes the variance matrix once multiplied by variance estimates

X0.scaled <- scaled.matrix[, 1:ncol(X0), drop = FALSE]
X1.scaled <- scaled.matrix[, ncol(scaled.matrix)]

# okay now we can set up the optimsiation problem
## the package gives the option to use `genoud` but we won't go into that
## for us we can set the initial parameter guesses as equal weight

SV1 <- rep(1/nvarsV,nvarsV)

# now we can estimate the parameters
res1 <- optimx(par=SV1, # initial guesses
                      fn=fn.V, # function
                      gr=NULL, # gradient not provided
                      hess=NULL, # donhessian not procided
                      method=optimxmethod,  
                      itnmax=NULL, # no max iterations
                      hessian=FALSE, # don't return hessian 
                             control=list(kkt=FALSE,
                                          starttests=FALSE,
                                          dowarn=FALSE,
                                          all.methods=all.methods),
                      X0.scaled = X0.scaled,
                      X1.scaled = X1.scaled,
                      Z0 = Z0,
                      Z1 = Z1,
                    quadopt = quadopt,
               margin.ipop = Margin.ipop,
               sigf.ipop = Sigf.ipop,
               bound.ipop = Bound.ipop)

# the collect.otpimx() is just a helpful little function that returns the parameters of itnerest for use
# run this code if you want to have a look at it
# view(collect.optimx)

res1 <- collect.optimx(res1,"min")

# round 2
Xall <- cbind(X1.scaled,X0.scaled)
Xall <- cbind(rep(1,ncol(Xall)),t(Xall))
Zall <- cbind(Z1,Z0)
Beta <- try(solve(t(Xall)%*%Xall)%*%t(Xall)%*%t(Zall),silent=TRUE) # this is just ols

Beta <- Beta[-1,]
V    <- Beta%*%t(Beta)
SV2  <- diag(V)
SV2 <- SV2 / sum(SV2)
  
res2 <- optimx(par=SV2, 
               fn=fn.V,
                             gr=NULL, hess=NULL, 
                             method=optimxmethod, itnmax=NULL, hessian=FALSE,
                             control=list(kkt=FALSE,
                                          starttests=FALSE,
                                          dowarn=FALSE,
                                          all.methods=all.methods),
                             X0.scaled = X0.scaled,
                             X1.scaled = X1.scaled,
                             Z0 = Z0,
                             Z1 = Z1,
                             quadopt = quadopt,
                             margin.ipop = Margin.ipop,
                             sigf.ipop = Sigf.ipop,
                             bound.ipop = Bound.ipop
                            )

res2 <- collect.optimx(res2,"min")

solution.v   <- abs(res2$par)/sum(abs(res2$par))

# great now that we have the solution for V* we can recover W(V*):

V <- diag(x=as.numeric(solution.v),nrow=nvarsV,ncol=nvarsV)
H <- t(X0.scaled) %*% V %*% (X0.scaled)
a <- X1.scaled
c <- -1*c(t(a) %*% V %*% (X0.scaled) )
A <- t(rep(1, length(c)))
b <- 1
l <- rep(0, length(c))
u <- rep(1, length(c))
r <- 0

res <- ipop(c = c, 
            H = H, 
            A = A, 
            b = b, 
            l = l, 
            u = u, 
            r = r,
            margin =  0.0005,
            maxiter = 1000,
            sigf = 5, 
            bound = 10)

solution.w <- as.matrix(primal(res))

# add names to out solutions
rownames(solution.w) <- colnames(X0)
colnames(solution.w) <- "w.weight"
names(solution.v) <- rownames(X0)

# estimate loss function values at solution
loss.w <- t(X1.scaled - X0.scaled %*% solution.w) %*% V %*% (X1.scaled - X0.scaled %*% solution.w)
loss.v <- t(Z1 - Z0 %*% as.matrix(solution.w)) %*% (Z1 - Z0 %*% as.matrix(solution.w)) /nrow(Z0)      

# now we have the optimised solutions for W(V)
#
data.frame(round(solution.w,4),
            output$names.and.numbers[-which(output$names.and.numbers[,2]==treatment.identifier),]
            )

matrix(round(solution.v,4),
       nrow=length(solution.v))
 

data.table(year = 1960:2005,
           co2_actual = Y1plot,
           co2_synth = Y0plot %*% solution.w,
           effects = Y1plot - (Y0plot %*% solution.w))


```



Now that we've explored what's going on under the hood, be thankful you can just call `Synth()` instead of having to code this up everytime! 

Big thanks to the author(s) for writing up this amazing package :)
