---
title: "synthetic_control_exploration"
output: 
  html_document:
    toc: true
    toc_float: true
---
<style>
/* Table of Contents Styles */
#TOC {
  color: #333333;
  font-family: 'Arial', sans-serif; 
  font-size: 16px; 
  border-color: #333333;
}

/* Header Styles */
#header {
  color: #B22222; 
  background-color: #F9F9F9; 
  opacity: 0.9; 
  font-family: 'Arial', sans-serif;
  font-size: 24px;
  padding: 15px; 
}

/* Body Styles */
body {
  color: #000000; 
  font-family: 'Arial', sans-serif;
  background-color: #FFFFFF; 
  line-height: 1.6; 
}

/*  Text Styles */
pre {
  color: #4D4D4D; 
  background-color: #F0F8FF; 
  border: 1px solid #B0C4DE; 
  padding: 10px; 
}

/* Header Styles */
h1, h2, h3 {
  color: #003366;
  font-family: 'Arial', sans-serif;
}

/* Link Styles */
a {
  color: #003366; 
  text-decoration: none; 
}

a:hover {
  text-decoration: underline; 
}
</style>

# Introduction

One really interesting estimator I never went through at university in any real depth was **Synthetic Control**. This is a great shame because it is a really exciting and honestly quite old estimator now *(and when I was studying!)*.

So I thought it might be nice to add a brief *(yet quite in depth)* exploration of this estimator! 

The most commonly used package in `R` is the `Synth` package by Jens Hainmueller and co: Jens Hainmueller aut cre, Alexis Diamond aut, Alberto Abadie aut (2011). “Synth: An R Package for Synthetic Control Methods in Comparative Case Studies.” _Journal of Statistical Software_, *42*(13), 1-17. <https://www.jstatsoft.org/v42/i13/>.

The package is really simple to use and generally robust for most quasi-experimental panel data applications. So let's explore what the Synthetic Control esitmator is, a neat application of it and then a bit of a peak under the hood as to what is going on in the `Synth()` function. 

```{r setup, message=FALSE}
library(data.table)
library(tidyverse)
library(Synth)
library(haven)
library(optimx)
library(kernlab)

options(scipen = 999)

```

# Synthetic Control Summary

... 

# Synthetic Control Example

To explore the estimator I will use the famous example from the paper: *Andersson, J.J., 2019. Carbon taxes and CO2 emissions: Sweden as a case study. American Economic Journal: Economic Policy, 11(4), pp.1-30*.

The paper exploits a natural experiment to ... 






```{r simple example}
co2 <- setDT(read_dta("/home/michael/Projects/essays/swedish_cp_replication/carbontax_fullsample_data.dta"))

#select countries and filter for years
co2_clean <- co2[country %in% c("Sweden",
                  "Australia",
                  "Belgium",
                  "Canada",
                  "Denmark",
                  "France",
                  "Greece",
                  "Iceland",
                  "Japan",
                  "New Zealand",
                  "Poland",
                  "Portugal",
                  "Spain",
                  "Switzerland",
                  "United States") & 
                   year > 1959 &
                   year < 2006]

# fix id
co2_clean[, Countryno := .GRP, by = .(country)]

# clean up environment
rm(co2)

# data prep object - we will explore all this intense looking stuff in depth later :)
D <- dataprep(foo = co2_clean,
              predictors = c("GDP_per_capita", 
                             "gas_cons_capita", 
                             "vehicles_capita", 
                             "urban_pop"),
              predictors.op = "mean",
              time.predictors.prior = 1980:1989,
              special.predictors = list(
                  list("CO2_transport_capita" , 1989 , "mean"),
                  list("CO2_transport_capita" , 1980 , "mean"),
                  list("CO2_transport_capita" , 1970 , "mean")
              ),
              dependent = "CO2_transport_capita",
              unit.variable = "Countryno",
              unit.names.variable = "country",
              time.variable = "year",
              treatment.identifier = 13,
              controls.identifier = c(1:12,14:15),
              time.optimize.ssr = 1960:1989,
              time.plot = 1960:2005)

# this is the wrapper function that calculates all the nice stuff for our synthetic control weights
synth_out <- synth(data.prep.obj = D, method = 'L-BFGS-B')

# fitted weight vectors (v for the predictors weights and w for the unit weights)
synth_out$solution.v
synth_out$solution.w

# simple effects calcs
effects <- D$Y1plot - (D$Y0plot %*% synth_out$solution.w)
sum(effects[32:length(effects)])
mean(effects[32:length(effects)])

# percentage changes
effects_prct <- (D$Y1plot - (D$Y0plot %*% synth_out$solution.w))/2.309894
sum(effects_prct[32:length(effects)])
mean(effects_prct[32:length(effects)])

# tables - creates some nice objects for reporting
synth_tables <- synth.tab(dataprep.res = D, synth.res = synth_out)

## table for weights w
synth_tables$tab.w

## special predictors (X matrices)
synth_tables$tab.pred
synth_tables$tab.v

# plots from ggplot
## path plot
data_plot <- data.table(year = 1960:2005,
                        co2_actual = D$Y1plot,
                        co2_synth = D$Y0plot %*% synth_out$solution.w,
                        effects = D$Y1plot - (D$Y0plot %*% synth_out$solution.w))

names(data_plot) <-  c("year", "co2_actual", "co2_synth", "effects")

data_plot %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = co2_actual, color = "Actual")) +
  geom_line(aes(y = co2_synth, color = "Synthetic")) +
  labs(title = "Path Plot of Actual vs Synthetic Outcomes",
       x = "Year",
       y = "Metric tons per capita (CO2 from transport)") +
  scale_color_manual(name = "Outcome", values = c("Actual" = "cornflowerblue", "Synthetic" = "blue")) +
  theme_classic()+
  theme(legend.position = "top")+
  geom_vline(xintercept = 1990, linetype = 'dashed', color = "black")


## gap plot
data_plot %>% 
  ggplot(aes(x = year, color)) +
  geom_line(aes(y = effects), color = "blue") +
  labs(title = "Gap plot of calculated effect",
       x = "Year",
       y = "Gap in metric tons per capita (CO2 from transport)") +
  theme_classic()+
  theme(legend.position = "top")+
  geom_hline(yintercept = 0, linetype = 'dashed', color = "black")+
  geom_vline(xintercept = 1990, linetype = 'dashed', color = "black")


# placebo test for loop
## initialise data
store <- matrix(NA,length(1960:2005),15)
colnames(store) <- unique(co2_clean$country)

# run loop
for(iter in 1:15){
 dataprep.out <-
              dataprep(foo = co2_clean,
                       predictors = c("GDP_per_capita" ,
                                      "gas_cons_capita" , 
                                      "vehicles_capita" ,
                                      "urban_pop") ,
                       predictors.op = "mean" ,
                       time.predictors.prior = 1980:1989 ,
                       special.predictors = list(
                         list("CO2_transport_capita" , 1989 , "mean"),
                         list("CO2_transport_capita" , 1980 , "mean") ,
                         list("CO2_transport_capita" , 1970 , "mean")
                         ),
                       dependent = "CO2_transport_capita",
                       unit.variable = "Countryno",
                       unit.names.variable = "country",
                       time.variable = "year",
                       treatment.identifier = iter, # note here the loop changes the treated unit for each loop
                       controls.identifier = c(1:15)[-iter],
                       time.optimize.ssr = 1960:1989,
                       time.plot = 1960:2005
                       )

# run synth
synth.out <- synth(data.prep.obj = dataprep.out, method = 'L-BFGS-B')

# store gaps
store[,iter] <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)
} # end of placebo loop

data <- store
rownames(data) <- 1960:2005

# Set bounds
gap.start     <- 1
gap.end       <- nrow(data)
years         <- 1960:2005
gap.end.pre  <- which(rownames(data)=="1989")

#  MSPE Pre-Treatment
mse  <- apply(data[ gap.start:gap.end.pre,]^2,2,mean)
sweden.mse <- as.numeric(mse[13])
# Exclude MSPE > 20 
data <- data[,mse<20*sweden.mse]
Cex.set <- 1

data_long <- data.table::melt(as.data.table(data, keep.rownames = TRUE),
                              id.vars = "rn",
                              variable.name = "Country",
                              value.name = "Gap")

data_long$Year <- as.integer(data_long$rn)

data_long %>% 
  ggplot(aes(x = Year, y = Gap, group = Country)) +
  geom_line(data = subset(data_long, Country != "Sweden"), 
            aes(color = "Control Countries"), color = "cornflowerblue", alpha = 0.4) +
  geom_line(data = subset(data_long, Country == "Sweden"), 
            aes(color = "Sweden"), color = "blue", linewidth = 1) +
  geom_vline(xintercept = 1990, linetype = "dotted", linewidth = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 1) +
  scale_x_continuous(breaks = seq(1960, 2005, 5)) +
  scale_y_continuous(limits = c(-0.5, 0.5)) +
  labs(x = "Year", y = "Gap in metric tons per capita (CO2 from transport)",
       title = "Gap in CO2 Emissions from Transport per Capita",
       subtitle = "Comparison between Sweden and Control Countries (Placebo)") +
  theme_classic() +
  theme(legend.position = "bottomleft") +
  annotate("text", x = 1983, y = -0.5, label = "Carbon Tax", size = 4) +
  annotate("segment", x = 1987, xend = 1989, y = -0.5, yend = -0.5, color = "black", arrow = arrow(length = unit(0.1, "inches"))) +
  guides(color = guide_legend(override.aes = list(linetype = c(1, 1), color = c("black", "gray"), size = c(1, 0.5))))

# Create a data frame for the weights
data.frame(country = synth_tables$tab.w$unit.names,
           Weight = synth_tables$tab.w$w.weights) %>% 
  ggplot(aes(x = reorder(country, -Weight), y = Weight)) +
  geom_bar(stat = "identity", fill = "cornflowerblue", color = "blue", alpha = 0.4) +
  labs(title = "Synthetic Control Unit Weights",
       x = "Countries",
       y = "Weights") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

rm(list=setdiff(ls(), "co2_clean"))

```

# A deep dive into Synthetic Control

Now that we've seen a simple application using `Synth` and the sort of results you can produce with it, lets dig into the code a bit deeper to get a better idea at what is going on underneath the hood! 



## data transformations
Firstly, let's look at that `dataprep` object and what it does:
- `foo` is the data.frame or table your data is stored in. In the above example this is the `co2_data`. 

```{r parameters}
# below are the key arguments used in estimating the synthetic control weights for the data prep object

# data prep objects parameters

# these are the predictors that form (part of the) X matrices used in the optimisation problem. This is a list of the column names as strings
predictors <-  c("GDP_per_capita", "gas_cons_capita", "vehicles_capita", "urban_pop") 

predictors.op <-  "mean" # the operation applie to the treated X matrix (i.e. mean or median etc - mean is usually sufficient) 

# these are the (special) predictors that form (part of the) X matrices used in the optimisation problem. This is a list of a list of the lags for the output variable we want to include.
special.predictors <-  list( 
                  list("CO2_transport_capita" , 1989 , "mean"),
                  list("CO2_transport_capita" , 1980 , "mean"),
                  list("CO2_transport_capita" , 1970 , "mean")
              ) 

# this is the variable of interest
dependent <-  "CO2_transport_capita"

# this is the column name identifying the unit dimension i.e. countries numebrs "Countryno"
# the code turns this into a column position index, so we can skip that part and just get set it as the first column 
unit.variable <-  1

# unit varibale names is the names we want to include in the final outputs i.e. the countries actual names and not just an index number
# the code also turns this into a column position index, so we can skip that part and just get set it as the 2nd column (country)
unit.names.variable <-  2

# the column that his the time dimension in our panel
# Similarly, the code turns this into a column position index, so we can skip that part and just get set it as the 3 column (year)
time.variable <-  3

# this is a arguement that identifies which of the `unit.variable` is the treated unit. For out case it is unit 13
treatment.identifier  <- 13

# this is a list identifying the remaining units that are controls (only one unit can be treated in this example)
controls.identifier <-  c(1:12,14:15)

# a list identifying the pretreatment periods over which the special predictors should be averaged
time.predictors.prior <-  1980:1989

# a list identifying the pretreatment periods over which the loss function is minimised
time.optimize.ssr <-  1960:1989

# this sets up the time we want to plot the data. Not important for our case tbh.
# if nothing is provided, this will default to `time.optimize.ssr`. But in our case we will set this to the whole period.
time.plot <-  1960:2005



```

Now that we've set up all the key arguements, let's explore the actual transformations.

```{r alt transformations}
# X0
X0 <- t(co2_clean[Countryno != treatment.identifier & year %in% time.predictors.prior, 
          .(GDP_per_capita = mean(GDP_per_capita, na.rm = TRUE),
            gas_cons_capita = mean(gas_cons_capita, na.rm = TRUE),
            vehicles_capita = mean(vehicles_capita, na.rm = TRUE),
            urban_pop = mean(urban_pop, na.rm = TRUE), 
            special.CO2_transport_capita.1989 = CO2_transport_capita[year == 1989],
            special.CO2_transport_capita.1980 = CO2_transport_capita[year == 1980]), 
          by = .(Countryno)] %>% select(-Countryno)) 


# have to add 1970 later because of time filter above
X0 <- rbind(X0, special.CO2_transport_capita.1970 = co2_clean[Countryno != treatment.identifier,CO2_transport_capita[year == 1970]])
# add columnn names
colnames(X0) <- controls.identifier

# X1
X1 <- t(co2_clean[Countryno == treatment.identifier & year %in% time.predictors.prior, 
          .(GDP_per_capita = mean(GDP_per_capita, na.rm = TRUE),
            gas_cons_capita = mean(gas_cons_capita, na.rm = TRUE),
            vehicles_capita = mean(vehicles_capita, na.rm = TRUE),
            urban_pop = mean(urban_pop, na.rm = TRUE), 
            special.CO2_transport_capita.1989 = CO2_transport_capita[year == 1989],
            special.CO2_transport_capita.1980 = CO2_transport_capita[year == 1980]), 
          by = .(Countryno)] %>% select(-Countryno)) 


# have to add 1970 later because of time filter above
X1 <- rbind(X1, special.CO2_transport_capita.1970 = co2_clean[Countryno == treatment.identifier,CO2_transport_capita[year == 1970]])
# add columnn names
colnames(X1) <- treatment.identifier

# Z0
Z0 <- dcast(co2_clean[Countryno != treatment.identifier & year %in% time.optimize.ssr,  
                      .(Countryno, year, CO2_transport_capita)], 
            year ~ Countryno, value.var = dependent) %>% select(-year) %>% as.matrix()

rownames(Z0) <- time.optimize.ssr

# Z1
Z1 <- dcast(co2_clean[Countryno == treatment.identifier & year %in% time.optimize.ssr,  
                      .(Countryno, year, CO2_transport_capita)], 
            year ~ Countryno, value.var = dependent) %>% select(-year) %>% as.matrix()

rownames(Z1) <- time.optimize.ssr

# plots 
# Y0plot
Y0plot <- dcast(co2_clean[Countryno != treatment.identifier,  
                      .(Countryno, year, CO2_transport_capita)], 
            year ~ Countryno, value.var = dependent) %>% select(-year) %>% as.matrix()

rownames(Y0plot) <- time.plot

# Y1plot
Y1plot <- dcast(co2_clean[Countryno == treatment.identifier,  
                      .(Countryno, year, CO2_transport_capita)], 
            year ~ Countryno, value.var = dependent) %>% select(-year) %>% as.matrix()

rownames(Y1plot) <- time.plot

# names and numbers
names.and.numbers <- data.frame(c(co2_clean[Countryno ==treatment.identifier, unique(country)],co2_clean[Countryno !=treatment.identifier, unique(country)]), 
                                c(co2_clean[Countryno ==treatment.identifier, unique(Countryno)],co2_clean[Countryno !=treatment.identifier, unique(Countryno)]))
names(names.and.numbers) <- c("unit.names", "unit.numbers")
```

There are a lot more checks and balances in the actual source code to make sure you don't mess anything up when you set up the `dataprep` object. But for our purposes of exploration, the above transformations show what the underlying source code achieves. So it will suffice for now :)

## Optimisation Problem

Okay great, we've now for all the data structures we need to estimate $W$. We have $Z_1$, $Z_0$, $X_1$ and $X_0$. We also have the matrices to calculate the counterfactual. But before, we do that we have to solve the optimisation problem! 


```{r alt optimisation}
# some arguements that are helpful to set up at the start
optimxmethod <- c("Nelder-Mead","BFGS")
# quadopt <-  "ipop"
# Margin.ipop <- 0.0005
# Sigf.ipop <- 5
# Bound.ipop <-  10
all.methods <- FALSE
# The first thing we do is normalise X by the variance - this returns some scaled matrices to us (scaled by variance of each predictor)
nvarsV <- dim(X0)[1]
big.dataframe <- cbind(X0, X1)
divisor <- sqrt(apply(big.dataframe, 1, var))

scaled.matrix <- t(t(big.dataframe) %*% ( 1/(divisor) * diag(rep(dim(big.dataframe)[1], 1)) )) # last bit is the identify matrix such that it becomes the variance matrix once multiplied by variance estimates

X0.scaled <- scaled.matrix[, 1:ncol(X0), drop = FALSE]
X1.scaled <- scaled.matrix[, ncol(scaled.matrix)]

# function to optimise
# So, let's explore how we can solve the optimisation problem. 
fn.V <- function (variables.v = stop("variables.v missing"), 
                  X0.scaled = stop("X0.scaled missing"),
                  X1.scaled = stop("X1.scaled missing"), 
                  Z0 = stop("Z0 missing"),
                  Z1 = stop("Z1 missing"), 
                  margin.ipop = 0.0005, 
                  sigf.ipop = 5,
                  bound.ipop = 10#, 
                  #quadopt = "ipop"
                  ){
  
  V <- diag(x = as.numeric(abs(variables.v)/sum(abs(variables.v))), nrow = length(variables.v), ncol = length(variables.v))
  H <- t(X0.scaled) %*% V %*% (X0.scaled)
  a <- X1.scaled
  c <- -1 * c(t(a) %*% V %*% (X0.scaled))
  A <- t(rep(1, length(c)))
  b <- 1
  l <- rep(0, length(c))
  u <- rep(1, length(c))
  r <- 0
  
  res <- ipop(c = c, 
              H = H, 
              A = A, 
              b = b, 
              l = l,
              u = u, 
              r = r, 
              bound = bound.ipop, 
              margin = margin.ipop, 
              maxiter = 1000, 
              sigf = sigf.ipop)
  
  solution.w <- as.matrix(primal(res))

  loss.w <- as.numeric(t(X1.scaled - X0.scaled %*% solution.w) %*% (V) %*% (X1.scaled - X0.scaled %*% solution.w))
  loss.v <- as.numeric(t(Z1 - Z0 %*% solution.w) %*% (Z1 -Z0 %*% solution.w))
  loss.v <- loss.v/nrow(Z0)
  return(invisible(loss.v))
}


# okay now we can set up the optimsiation problem
## the package gives the option to use `genoud` but we won't go into that
## for us we can set the initial parameter guesses as equal weight

SV1 <- rep(1/nvarsV,nvarsV)

# now we can estimate the parameters
opt_res <- optimx::optimx(par=SV1, # initial guesses
                  fn=fn.V, # function
                  gr=NULL, # gradient not provided
                  hess=NULL, # donhessian not procided
                  method=optimxmethod,
                  itnmax=NULL, # no max iterations
                  hessian=FALSE, # don't return hessian
                  control=list(kkt=FALSE,
                               starttests=FALSE,
                               dowarn=FALSE,
                               all.methods=all.methods),
                  X0.scaled = X0.scaled,
                  X1.scaled = X1.scaled,
                  Z0 = Z0,
                  Z1 = Z1#,
                  # quadopt = quadopt,
                  # margin.ipop = Margin.ipop,
                  # sigf.ipop = Sigf.ipop,
                  # bound.ipop = Bound.ipop
                  )

# the collect.otpimx() is just a helpful little function that returns the parameters of itnerest for use
# run this code if you want to have a look at it
# view(collect.optimx)

opt_res <- collect.optimx(opt_res,"min")

solution.v   <- abs(opt_res$par)/sum(abs(opt_res$par))

# great now that we have the solution for V* we can recover W(V*):

V <- diag(x=as.numeric(solution.v),nrow=nvarsV,ncol=nvarsV) # solution for V act as weights
H <- t(X0.scaled) %*% V %*% (X0.scaled) 
a <- X1.scaled 
c <- -1*c(t(a) %*% V %*% (X0.scaled) ) # vector appearing in quadratic function
A <- t(rep(1, length(c))) # defines constaints for quadratic optimisation
b <- 1 # optimised vector constraint (upper)
l <- rep(0, length(c)) # lwoer bound
u <- rep(1, length(c)) # upper bound
r <- 0 # optimised vector constraint (lower)  

# https://www.rdocumentation.org/packages/kernlab/versions/0.9-29/topics/ipop
# https://en.wikipedia.org/wiki/Quadratic_programming
quad_res <- kernlab::ipop(c = c, 
            H = H, 
            A = A, 
            b = b, 
            l = l, 
            u = u, 
            r = r,
            margin =  0.0005, # how close we can get to the constraints
            maxiter = 1000, # max number of iterations
            sigf = 5, # precision
            bound = 10) # clipping bound for variables

solution.w <- as.matrix(primal(quad_res))

# add names to out solutions
rownames(solution.w) <- colnames(X0)
colnames(solution.w) <- "w.weight"
names(solution.v) <- rownames(X0)

# estimate loss function values at solution
loss.w <- t(X1.scaled - X0.scaled %*% solution.w) %*% V %*% (X1.scaled - X0.scaled %*% solution.w)
loss.v <- t(Z1 - Z0 %*% as.matrix(solution.w)) %*% (Z1 - Z0 %*% as.matrix(solution.w)) /nrow(Z0)      

# now we have the optimised solutions for W(V)
# table for W
data.frame(round(solution.w,4),
           names.and.numbers[-which(names.and.numbers[,2]==treatment.identifier),])

# predictor comparisons
tab.pred <- round(cbind(X1,
                        as.matrix(X0)%*%as.matrix(solution.w),
                        apply(as.matrix(X0),1,mean)), 4)
                          
colnames(tab.pred) <- c("Treated", "Synthetic", "Sample Mean")

tab.pred

# table for V
tab.v <- matrix(round(solution.v, 4), nrow=length(solution.v))
rownames(tab.v) <- rownames(X1)
colnames(tab.v) <- c("v.weights")

# plot output
data_plot2 <-  data.table(year = 1960:2005,
           co2_actual = Y1plot,
           co2_synth = Y0plot %*% solution.w,
           effects = Y1plot - (Y0plot %*% solution.w))

names(data_plot2) <-  c("year", "co2_actual", "co2_synth", "effects")

data_plot2 %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = co2_actual, color = "Actual")) +
  geom_line(aes(y = co2_synth, color = "Synthetic")) +
  labs(title = "Path Plot of Actual vs Synthetic Outcomes",
       x = "Year",
       y = "Metric tons per capita (CO2 from transport)") +
  scale_color_manual(name = "Outcome", values = c("Actual" = "cornflowerblue", "Synthetic" = "blue")) +
  theme_classic()+
  theme(legend.position = "top")+
  geom_vline(xintercept = 1990, linetype = 'dashed', color = "black")
```


Now that we've explored what's going on under the hood, be thankful you can just call `Synth()` instead of having to code this up everytime! 

Big thanks to the author(s) for writing up this amazing package :)
